---
title: "Distractor Rejection Data Set Filtering"
output:
  html_document:
    theme: cosmo
    highlight: zenburn
  pdf_document: default
---

# Import libraries

```{r, message=FALSE}
source("funs.R")
library(ggpubr)
```

# Read Data from E-Prime

One caveat: eMerge adds an export location as a first line in the merged data set. 
Delete that line manually or else R will read it in incorrectly.

```{r, step1.1}
behavioral <- read_csv("disrej_color_edat.csv")
behavioral
```

## E-Prime Column Names

Variable names in the original file contain square brackets, periods, and capital 
letters which can make the data impossible to work with, so we have to rename the
columns. This one only replaces periods and brackets with underscores which is not
perfect but is enough to make the data analyzable.

```{r, step1.2}
behavioral <- behavioral %>%
  rename_with(~sub('[.]+', '_', .), perl=TRUE) %>%
  rename_with(~sub('\\[', '_', .), perl=TRUE) %>%
  rename_with(~sub('\\]', '_', .), perl=TRUE) %>%
  rename_with(~sub('_$', '', .), perl=TRUE) %>%
  filter(Running_Block == "BlockList") 
```
## E-Prime Data Filtering

Now filter out all blocks that are not 'BlockList' so
that we are using the data only from the real trials. Then, it will create a
tibble, where 'Subject' is the subject number and 'accuracy' is their accuracy 
for all trials in percent. The best way is to just eyeball it and see where the 
accuracy is lower than 70%, then remember the subject numbers for the next step.


```{r, step1.3}
behavioral %>%
  group_by(Subject) %>%
  summarize(accuracy = (sum(Query_ACC_Trial) / n() * 100))
```

# Read Data from EyeLink

Import the Interest Area Report.

```{r, step2.1}
ia <- read_csv("disrej_color_data.csv")
ia
```

## Remove Specific Subjects

In this step, the data from specific subject numbers is removed. Subject numbers
from the accuracy calculation step go in here. Also, if the data was not recorded
for some participants, it's also removed in here. From Distractor Rejection Color
experiment, participant 6 did not have a part of the block recorded, so I will
remove it here from the eye movement data set.

-- I commented out the code that removes 6 since invalid trials will be removed
with the filters below.

```{r, step2.2}
ia <- ia %>%
  group_by(Subject_Num) #%>%
  #filter(Subject_Num != c(6))
ia
```

## Clean eye movement data

Looking at IA label and creating a new column (is_target) that will note whether
an IA is a distractor or a target. Also, putting the information about whether
an interest are was fixated on into the 'did_fixate' column.

```{r, step2.3}
ia <- ia %>%
  mutate(is_target = "default") %>%
  mutate(is_target = if_else(startsWith(IA_LABEL, "T"), "Target", is_target)) %>%
  mutate(is_target = if_else(startsWith(IA_LABEL, "D"), "Distractor", is_target)) %>%
  mutate(did_fixate = "default") %>%
  mutate(did_fixate = if_else(IA_FIRST_RUN_FIXATION_COUNT %in% c("."), 0, 1))
ia
```

## Outlier Removal

Calculating mean +- SD for the whole data set and then comparing each RT such that
it should be between 2.5 mean-SD and 200.

```{r, step2.4}
sd_above <- mean(ia$VS_RT) + (2.5 * sd(ia$VS_RT))

ia <- ia %>%
  group_by(Subject_Num, TrialNum) %>%
  filter(between(VS_RT, 200, sd_above))
ia
```

## More Filtering

This step should include RT between 200 and 2.5SD, did_fixate = 1, is_target = "Target".
I did not include the accuracy filter since this data is in the other table, so it 
will be filtered after a table join. I also excluded Block_Num 'UNDEFINED' since
it refers to the practice trials.

```{r, step2.5}
ia <- ia %>%
  filter(is_target == "Target") %>%
  filter(did_fixate == 1) %>%
  filter(Block_Num != "UNDEFINED")
ia
```

## Join tables

Behavioral and the IA table contain the data that are carved up differently (by-trial and by-IA),
so it has to be joined by the subject number, block number, and the trial number to display correctly.
Since there are almost 300 features in this data set, it makes sense to leave only
the important ones. Also, I changed the data type of Block_Num back to numeric as
it was 'chr' because of some values being UNDEFINED. Also the accuracy filter is 
applied here.

Another modification here is the 1101ms correction applied to the ia_first_fixation_time.
The accuracy filter is applied here as well as the ttp > 50 and ia_first_fixation_time > 0.


```{r, step2.6}
ia <- ia %>%
  mutate(Block_Num = as.numeric(Block_Num)) %>%
  left_join(behavioral, by = c("Subject_Num" = "Subject", "TrialNum" = "TrialNum_Trial", "Block_Num" = "BlockNum_Block")) %>%
  filter(Query_ACC_Trial == 1) %>%
  select(Subject_Num, Block_Num, TrialNum, DistType, VS_RT, IA_FIRST_FIXATION_INDEX,IA_FIRST_FIXATION_RUN_INDEX, IA_FIRST_FIXATION_VISITED_IA_COUNT, IA_FIRST_FIXATION_TIME, IA_DWELL_TIME, VSTarget_Trial, INTEREST_AREA_FIXATION_SEQUENCE) %>%
  mutate(IA_FIRST_FIXATION_INDEX = as.numeric(IA_FIRST_FIXATION_INDEX)) %>%
  mutate(IA_FIRST_FIXATION_RUN_INDEX = as.numeric(IA_FIRST_FIXATION_RUN_INDEX)) %>%
  mutate(IA_FIRST_FIXATION_VISITED_IA_COUNT = as.numeric(IA_FIRST_FIXATION_VISITED_IA_COUNT)) %>%
  mutate(IA_FIRST_FIXATION_TIME = as.numeric(IA_FIRST_FIXATION_TIME) - 1101) %>%
  mutate(IA_DWELL_TIME = as.numeric(IA_DWELL_TIME)) %>%
  mutate(ttp = VS_RT - IA_FIRST_FIXATION_TIME) %>%
  filter(ttp > 50) %>%
  filter(IA_FIRST_FIXATION_TIME > 0)

colnames(ia) <- stri_trans_tolower(names(ia)) # make the col-names lowercase
ia
```

## Categorize

Read in the object category data set that has the object category matched with
the file names and the mds category table that specifies the MDS values for each 
category. Join both tables with the main data set.

```{r, step 2.7}
obj_cat <- read_csv("categories.csv")
mds <- read_csv("mds_categories.csv")

ia <- ia %>%
  left_join(obj_cat, by = c("vstarget_trial" = "filename")) %>%
  left_join(mds, by = c("category")) %>%
  select(subject_num, block_num, trialnum, disttype, vs_rt, ia_first_fixation_index, ia_first_fixation_run_index, ia_first_fixation_visited_ia_count, ia_first_fixation_time, ia_dwell_time, ttp, category, mds_category, interest_area_fixation_sequence)
  
ia
```

## Prepare and Export

Take each DV and and create a table for the analyses that will be written into 
separate CSV files.

```{r, step 2.8}

ia %>%
  ungroup() %>%
  select(subject_num, block_num, trialnum, disttype, mds_category, vs_rt) %>%
  pivot_wider(names_from = c(disttype, mds_category), values_from = vs_rt) %>%
  group_by(subject_num) %>%
  summarize(close_high = mean(Close_high, na.rm = TRUE), close_low = mean(Close_low, na.rm = TRUE), far_low = mean(Far_low, na.rm = TRUE), far_high = mean(Far_high, na.rm = TRUE)) %>%
  write.csv("analysis_vsrt.csv")

ia %>%
  ungroup() %>%
  select(subject_num, block_num, trialnum, disttype, mds_category, ia_first_fixation_index) %>%
  pivot_wider(names_from = c(disttype, mds_category), values_from = ia_first_fixation_index) %>%
  group_by(subject_num) %>%
  summarize(close_high = mean(Close_high, na.rm = TRUE), close_low = mean(Close_low, na.rm = TRUE), far_low = mean(Far_low, na.rm = TRUE), far_high = mean(Far_high, na.rm = TRUE)) %>%
  write.csv("analysis_ia_first_fixation_index.csv")

ia %>%
  ungroup() %>%
  select(subject_num, block_num, trialnum, disttype, mds_category, ia_first_fixation_run_index) %>%
  pivot_wider(names_from = c(disttype, mds_category), values_from = ia_first_fixation_run_index) %>%
  group_by(subject_num) %>%
  summarize(close_high = mean(Close_high, na.rm = TRUE), close_low = mean(Close_low, na.rm = TRUE), far_low = mean(Far_low, na.rm = TRUE), far_high = mean(Far_high, na.rm = TRUE)) %>%
  write.csv("analysis_ia_first_fixation_run_index.csv")

ia %>%
  ungroup() %>%
  select(subject_num, block_num, trialnum, disttype, mds_category, ia_first_fixation_visited_ia_count) %>%
  pivot_wider(names_from = c(disttype, mds_category), values_from = ia_first_fixation_visited_ia_count) %>%
  group_by(subject_num) %>%
  summarize(close_high = mean(Close_high, na.rm = TRUE), close_low = mean(Close_low, na.rm = TRUE), far_low = mean(Far_low, na.rm = TRUE), far_high = mean(Far_high, na.rm = TRUE)) %>%
  write.csv("analysis_ia_first_fixation_visited_ia_count.csv")

ia %>%
  ungroup() %>%
  select(subject_num, block_num, trialnum, disttype, mds_category, ia_first_fixation_time) %>%
  pivot_wider(names_from = c(disttype, mds_category), values_from = ia_first_fixation_time) %>%
  group_by(subject_num) %>%
  summarize(close_high = mean(Close_high, na.rm = TRUE), close_low = mean(Close_low, na.rm = TRUE), far_low = mean(Far_low, na.rm = TRUE), far_high = mean(Far_high, na.rm = TRUE)) %>%
  write.csv("analysis_ia_first_fixation_time.csv")

ia %>%
  ungroup() %>%
  select(subject_num, block_num, trialnum, disttype, mds_category, ia_dwell_time) %>%
  pivot_wider(names_from = c(disttype, mds_category), values_from = ia_dwell_time) %>%
  group_by(subject_num) %>%
  summarize(close_high = mean(Close_high, na.rm = TRUE), close_low = mean(Close_low, na.rm = TRUE), far_low = mean(Far_low, na.rm = TRUE), far_high = mean(Far_high, na.rm = TRUE)) %>%
  write.csv("analysis_ia_dwell_time.csv")

ia %>%
  ungroup() %>%
  select(subject_num, block_num, trialnum, disttype, mds_category, ttp) %>%
  pivot_wider(names_from = c(disttype, mds_category), values_from = ttp) %>%
  group_by(subject_num) %>%
  summarize(close_high = mean(Close_high, na.rm = TRUE), close_low = mean(Close_low, na.rm = TRUE), far_low = mean(Far_low, na.rm = TRUE), far_high = mean(Far_high, na.rm = TRUE)) %>%
  write.csv("analysis_ttp.csv")
```